Visual Diagnostics of a Model Explainer -- Tools for the Assessment of LIME Explanations

### 2020-06-04 paper.Rnw
edits through section 3.0 (almost)

In the field of statistics, there are two main uses for models: inference and prediction. Machine learning models are often used for the latter purpose. These models have proven to perform well in a wide range of prediction problems, but the accuracy of many machine learning models comes at the cost of interpretability due to their algorithmic complexity (hence the phrase "black-box models"). Model interpretability allows for the assessment and understanding of how the model produces predictions. In application, the ability to assess and understand model is crucial in important and crutial in areas areas with high stakes decisions such as health care and forensics science. The increased use of machine learning models and the introduction of the General Data Protection Regulation (GDPR) in 2018 \citep{goodman:2016} has resulted in a dramatic increase in research in the area of explainable machine learning that focuses on developing ways to explain output from machine learning algorithms.

Throughout this paper, we will make a distinction between interpretability and explanability of models. We define {\it interpretability} as the ability to directly use model parameters to understand the relationships in the data captured by the model: e.g., a linear model coefficient associated with a predictor variable indicates the amount the response variable changes based on a change in the predictor variable. In contrast, {\it explanability} is defined as the ability to use the model in an indirect manner to understand the relationships in the data captured by the model: e.g., partial dependence plots depict the marginal relationship between model predictions and predictor variables \citep{friedman:2001}.

Numerous methods have been proposed to provide explanations for black-box model predictions \citep{guidotti:2018, mohseni:2018, molnar:2019}. Some are specific to one type of model (e.g. \citep{bau:2017} and \citep{hara:2016}), and others are model-agnostic (e.g. \citep{greenwell:2018} and \citep{strumbelj:2014}). In this paper, we focus on the model-agnostic method of LIME (local interpretable model-agnostic explanations) \citep{ribeiro:2016}.

@@

@
\caption{A conceptual depiction of an unfaithful local explainer model in the immediate neighborhood of a prediction of interest. A ridge regression model is fit to the same black-box model predictions and predictor variables as in \autoref{fig:concept-plot-good}, but the weights are computed using a Gower exponent of 1. The explainer model provides a global approximation of the complex model in regards to the prediction of interest.}
\label{fig:concept-plot-bad}
\end{figure*}

\subsection{LIME}

LIME falls under the category of model explainers \citep{ribeiro:2016} \citep{mohseni:2018}. A model explainer is a method that uses a surrogate model to relate the predictor variables to the black-box model predictions. We will distinguish between the terms of model explainer and explainer model: by \emph{model explainer} we denote the method for explaining a complex model using a surrogate model, while the \emph{explainer model}, or simply the \emph{explainer}, is the surrogate model.

While some model explainers are focused on understanding a model at the global level, LIME claims to provide local explanations. Additionally, LIME is designed to work with any model and to produce easily understandable explanations \citep{ribeiro:2016}. Conceptually, LIME fits a simple interpretable model, the explainer model,  meant to capture the behavior of the (complex) black-box model in a local region around a prediction of interest. The  simple model then provides interpretable estimates for variables that most influenced the prediction made by the complex model.

\autoref{fig:concept-plot-good} provides a visualization of this conceptual understanding of LIME. The two plots show the predictions from a hypothetical black-box model plotted against the two predictor variables used to fit the hypothetical model. The diamond shaped points represent a prediction of interest. A Gower distance metric \citep{gower:1971} is used to define locality: the size of the points represent the inverse of the Gower distance as a reflection of the proximity to the prediction of interest. Here, an exponent of \Sexpr{good_gower_power} is used to emphasize interest in a very local region around the point of interest. A ridge regression model weighted by the proximity values is used as the explainer model. The black-box predictions are used as the response variable, standardized versions of Features 1 and 2 are included as predictor variables. Standardized features allow direct comparisons of the model coefficients. The explainer model is depicted by the black lines in the figure.

The plot on the left of \autoref{fig:concept-plot-good} shows the relationship between the black-box predictions and Feature 1. Here, the explainer model is plotted with Feature 2 fixed at the observed value of Feature 2 for the prediction of interest. The explainer model captures the relationship in the immediate neighborhood around the prediction of interest  with a slope of \Sexpr{coefs_data %>% filter(case == "good", feature == "Feature 1") %>% pull(slope) %>% round(3)}. The plot on the right shows that there is no global or local relationship between the black-box predictions and Feature 2. Here, the explainer model is plotted with Feature 1 set to be the observed value of Feature 1 for the prediction of interest, and it has a small slope of \Sexpr{coefs_data %>% filter(case == "good", feature == "Feature 2") %>% pull(slope) %>% round(3)}. The magnitude of the slope associated with Feature 1 is larger than the slope of Feature 2, resulting LIME to report that Feature 1 plays a more important role in the prediction made by the black-box model.

@@

Explainer models are meant to provide clarity to the process of predictive modeling, but since the explainer model is an approximation of the complex model and not a direct interpretation, explanations produced by an explainer model are subject to the quality of the approximation. Thus, explainer models add another layer of complexity to predictive models by requiring yet another model that needs to be assessed.

The choice of exponent in the scenario depicted in \autoref{fig:concept-plot-good} is crucial. High values ensure an up-weight of observations closer to the prediction of interest. Consider the example depicted in \autoref{fig:concept-plot-bad}. The plots show the same data as \autoref{fig:concept-plot-good}, but the Gower distance metric exponent was decreased to \Sexpr{bad_gower_power} (the default exponented in the \emph{lime} R package). This causes the observations that are further away from the prediction of interest to be given larger weights than before. In \autoref{fig:concept-plot-good}, the explainer model captures the relationship between the black-box predictions in the immediate neighborhood of the prediction of interest. This cannot be said of the explainer model in \autoref{fig:concept-plot-bad}. In addition, the magnitude of the slope (\Sexpr{coefs_data %>% filter(case == "bad", feature == "Feature 2") %>% pull(slope) %>% round(3) %>% abs()}) associated with Feature 2 is larger than that of Feature 1 (\Sexpr{coefs_data %>% filter(case == "bad", feature == "Feature 1") %>% pull(slope) %>% round(3) %>% abs()}). Thus, this explainer model actually provides a misleading explanation that Feature 2 plays a more important role in predicting the observation of interest.

There are various ways in which LIME can fail. One of the biggest difficulties with LIME is determining how to specify a local region \citep{molnar:2019} due to an unclear definition of what a local region means and how to apply LIME to achieve an appropriate local region. \autoref{fig:concept-plot-bad} demonstrates an explainer model that provides misleading explanations due to poorly specified weights. \citet{laugel:2018} show additional examples of undesirable explanations produced by LIME when the local region was chosen inadequately. \citet{alvarezmelis:2018} raise another concern pertaining to the robustness of explanations from LIME and other post-hoc explainers: they find that even small changes in predictor variables can lead to very different LIME explanations. Additionally, LIME relies on a good approximation of the explainer model to the complex model. \citet{ribeiro:2016} acknowledge that ``if the underlying model is highly non-linear even in the locality of the prediction, there may not be a faithful explanation" with a linear regression model as the explainer model.

As a result of the various ways LIME explanations can fail, it is important to assess LIME explanations. In this paper, we lay out the set of claims about LIME made by \citet{ribeiro:2016} and propose four visualizations for the assessment of these claims: (1) \emph{explanation scatterplot}, (2) \emph{prediction comparison plot}, (3) \emph{feature heatmap}, (4) \emph{assessment metric plot}.

While LIME has been implemented for image, tabular, and text data, we will only focus on tabular data. For additional simplicity, we will only discuss classification prediction models with a dichotomous response variable and continuous predictor variables. However, the proposed diagnostics may be extended to a wider range of situations.

\subsection{Overview of Paper}

The reminder of the paper is structured as follows. Section \ref{background} provides background and the claims made by \citet{ribeiro:2016} about LIME. In Section \ref{diagnostics}, we introduce the suggested diagnostic plots. Then in Section \ref{application}, we demonstrate the use of the diagnostics to assess LIME explanations for a random forest model fit to a forensics bullet matching dataset. Section \ref{discussion} concludes with a discussion on extenstion and limitations of the diagnostic plots and concerns about LIME in regards to the claims made by \citet{ribeiro:2016} brought about by the visualization examples in this paper that agree with \citet{alvarezmelis:2018}, \citet{laugel:2018}, and \citet{molnar:2019}.

@@

LIME was introduced in 2016 by \citet{ribeiro:2016}. LIME is presented as a model explainer that produces easily understandable explanations for individual predictions from any predictive model \citep{ribeiro:2016}. As described by \citet{laugel:2018}, the general form of the LIME algorithm can be divided into three steps:

@@

\item \emph{Explainer Model Fitting}: Apply the black-box model to the simulated data to obtain predictions. Compute the distance between each of the simulated observations and the prediction of interest. Fit an interpretable model with the black-box predictions from the simulated data as the response, the transformed simulated data as the predictors, and the distances as weights. This model is the explainer model.

@@

\citet{ribeiro:2016} make the following set of claims regarding the performance of LIME:

@@

LIME was initially implemented in a Python package\footnote{https://github.com/marcotcr/lime} by the original authors. The adaption to an R package is by Thomas Lin-Pedersen\footnote{https://github.com/thomasp85/lime}. In this paper, we use a forked version\footnote{https://github.com/goodekat/lime} of the development version of the R package \emph{lime} (\Sexpr{packageVersion("lime")}) \citep{pedersen:2020}. The forked version of \emph{lime} allows us to export internal values of the procedure not already exported by \emph{lime}.

@@

At a high level, the concept of LIME is relatively simple: an interpretable model is used to approximate a complex model in a local region. In practice, the implementation of LIME is not straightforward, and there is research is being done to improve the procedure \citep{laugel:2018}. As described in \autoref{background}, there are various parameter choices that have to be made when applying LIME: the simulation method, the feature selection method, and how weights are calculated.  We will refer to these choices as tuning parameters. Without an assessment of the explainer model, it is not clear if the tuning parameters selected when applying LIME resulted in a useful explanation and LIME becomes a black-box procedure of its own. To counteract the black-box nature, we suggest the use of visualizations of each step of the LIME algorithm. In this section, we introduce four suggested visual diagnostic plots for the assessment of LIME. The diagnostic plots focus on assessing the LIME claims at different levels of application:

\begin{enumerate}
\item Explanation scatterplot (Section \ref{plot1}): Comparison of the explainer and complex models for an individual prediction of interest.
\item Prediction comparison plot (Section \ref{plot1}): Comparison of explainer and complex model predictions for a set of predictions.
\item Feature Heatmap: Comparison of features selected by LIME across applications of LIME with different tuning parameters.k
\item Assessment metric plot: Comparison of performance metrics for LIME across applications of LIME with different tuning parameters. 
\end{enumerate}

@@

The dataset is divided into a training set of \Sexpr{dim(sine_data_train)[1]} observations and a testing set of \Sexpr{dim(sine_data_test)[1]} observations. A random forest model is fit using the R package \emph{randomForest} (version \Sexpr{packageVersion("randomForest")}) \citep{liaw:2002} with the default settings. The model is applied to the testing set to obtain predictions. \autoref{fig:sine-plot-data} shows scatterplots of $x_2$ versus $x_1$ from the training data (left) and the testing data (right). Both plots include the true classification boundary of the rotated sine function plotted as the solid black line. The training data are colored by the observed response variable ($y$), and the testing data are colored by the prediction probabilities from the random forest model ($\hat{y}$).

The random forest model misclassifies \Sexpr{dim(sine_data_test %>% filter(y != rfpred))[1]} points on the classification boundary. These are identified by circles in \autoref{fig:sine-plot-data}. For the diagnostic plot presented in Section \ref{plot1}, we focus on one of these points: the observation with $(x_1, x_2, x_3)$ coordinates of (\Sexpr{round(sine_poi$x1, 2)}, \Sexpr{round(sine_poi$x2, 2)}, \Sexpr{round(sine_poi$x3, 2)}) indicated by a diamond in \autoref{fig:sine-plot-data}.

@@

We apply LIME to the \data \ using \emph{lime} with the default settings. The user is asked to specify the number of features shown in the explanations. Here, we set this value to two, because we expect that $x_1$ and $x_2$ will be the main drivers in the prediction. Note that the complexity of an explanation goes hand in hand with the number of features selected to be included in the explanation. Setting the number of features to a small number will make the explanation simpler but not necessarily more correct.

To demonstrate the proposed diagnostic plots, we applied six different implementations of LIME to the sine data to select the top two features associated with each of the random forest predictions in the testing dataset. This was done using six different sets of tuning parameters via \sout{our} R package \emph{limeaid} \footnote{Available on GitHub at https://github.com/goodekat/limeaid}. Quantile bins were used for five of the implementations, and the number of bins for these implementations was varied from 2 to 6. The sixth implementation of LIME used the kernel density estimates. The feature selection method and the computation of the weights were set to the default options in \emph{lime} of forward selection and Gower distance raised to an exponent of 1.

@@

\kgc{Add intro paragraph and edit text as necessary in this subsection}

@@

\subsection{Feature Heatmap Diagnostic Plot} \label{feature-heatmap}

@@

\subsection{Assessment Metric Plot}

@@

\kgc{Break into sections: (1) visualization limitations (2) future research (3) concerns about LIME}

%\kgc{Possible content to add here:} \kgc{Need to work this in}: \hh{In terms of limitations of the visualizations: first two features span the x and y axis, additional features can be used for faceting. four features is completely feasible, beyond that the visualization might not work as well. The number of bins should be in the single digits. A large number of small multiples \citep{Becker:1996gy} needs a strong signal in the data to be effective - otherwise the visualization might be affected by change blindness \citep{simonsChangeBlindness1997, review19}. }

@@

%In the case of faceting, the} \hh{first two features} \kge{can} \hh{span the x and y axis,} \kge{and} \hh{additional features can be} \kge{used as the facets.} We have included some examples of such plots in Appendix \ref{scatter-plots} that take into the feature $x_3$ from the \data \ in the visualization of the LIME procedure.

%\hh{\sout{In terms of limitations of the visualizations: first two features span the x and y axis, additional features can be used for faceting. four features is completely feasible, beyond that the visualization might not work as well. The number of bins should be in the single digits. A large number of small multiples (reference to Cleveland?) needs a strong signal in the data to be effective - otherwise the visualization might be affected by change blindness} \citet{simons:1997, vanderplas:2020}.}

@@

\kgc{Need to add a discussion on using our visualizations in situations without tabular data or a dichotomous response variable.}\\

In this paper, we have presented visualizations that depict the steps in the LIME algorithm (\autoref{procedure}) and diagnose LIME explanations (Sections \ref{set-visuals} and \ref{implementation-visuals}). Our intentions with these visualizations were to provide insight on how LIME works, assess the ability of the explainer model to capture the complex predictive model, and compare LIME explanations produced by different tuning parameters. While the visualizations accomplish these tasks, they also expose examples of the failings of LIME. To address the discovered failings of LIME, we will reconsider each of the claims about the performance of LIME made by \citep{ribeiro:2016} (interpretability, faithfulness, linearity, and localness) in light of the insights gained from the diagnostic visualizations.

@@

\kgc{**Need to rework this paragraph and the next two. The ideas are here, but they are worded very poorly.} Even though an explanation will be interpretable as long as the complexity of the explainer model is appropriately chosen, the presentation of a LIME explanation in the key visual used by \citep{ribeiro:2016} (shown in \autoref{fig:sine-plot-poi}) could lead to an over simplified interpretation of the explainer model. Without a solid understanding of the details behind the fitting of the explainer model, the deeper meaning of the bars in the figure is lost.

@@

\kgc{Need to work this into the discussion: We acknowledge that not all prediction problems will produce trends as easy to interpret as the visualizations we have shown in the section, but we still believe that visualizations of the procedure can provide insight into the quality of the LIME explanation. We hope these figures will provide a starting point (if not an ending point) for the assessment of a single LIME explanation. Perhaps the inability to interpret a plot is itself telling, or the act of creating extensions of these visualizations to work with a different scenarios will lead to an understanding of the trustworthiness of the explanation.}

@@

>>>>

In the field of statistics, there are two main uses for models: inference and prediction. Machine learning models are often used for the latter purpose. These models have proven to perform well in a wide range of prediction problems, but the accuracy of many machine learning models comes at the cost of interpretability due to their algorithmic complexity (hence the phrase "black-box models"). Model interpretability allows for the assessment and understanding of how the model produces predictions. The lack of the ability to assess and understand a model makes it difficult to trust the model, especially in areas areas with high stakes decisions such as health care and forensics science. The increased use of machine learning models in applications and the introduction of the General Data Protection Regulation (GDPR) in 2018 \citep{goodman:2016} has resulted in a dramatic increase in research in an area known as explainable machine learning, which focuses on developing ways to explain output from machine learning algorithms.

Throughout this paper, we make a distinction between interpretability and explanability of models. We define {\it interpretability} as the ability to directly use model parameters to understand the relationships in the data captured by the model: e.g., a linear model coefficient associated with a predictor variable indicates the amount the response variable changes based on a change in the predictor variable. In contrast, {\it explanability} is defined as the ability to use the model in an indirect manner to understand the relationships in the data captured by the model: e.g., partial dependence plots depict the marginal relationship between model predictions and predictor variables \citep{friedman:2001}.

Numerous methods have been proposed to provide explanations for black-box model predictions \citep{guidotti:2018, mohseni:2018, molnar:2019}. Some are specific to one type of model (e.g. \citep{bau:2017} and \citep{hara:2016}), and others are model-agnostic (e.g. \citep{greenwell:2018} and \citep{strumbelj:2014}). In this paper, we focus on the model-agnostic method of LIME \citep{ribeiro:2016}.

@@

@
\caption{A conceptual depiction of an unfaithful local explainer model in the immediate neighborhood of a prediction of interest. A ridge regression model is fit to the same black-box model predictions and predictor variables as in \autoref{fig:concept-plot-good}, but the weights are computed using a Gower exponent of 1. The explainer model is unfaithful to the complex model in the immediate neighborhood of the prediction of interest.}
\label{fig:concept-plot-bad}
\end{figure*}

\subsection{Conceptual Description of LIME}

LIME (local interpretable model-agnostic explanations) is a method that uses a surrogate model to relate predictor variables to black-box model predictions (i.e. a model explainer) \citep{ribeiro:2016}. We distinguish between the terms of model explainer and explainer model: by \emph{model explainer} we denote the method for explaining a complex model using a surrogate model, while the \emph{explainer model}, or simply the \emph{explainer}, is the surrogate model.

While some model explainers are focused on understanding a model at the global level, LIME claims to provide explanations for individual predictions (local). Additionally, LIME is designed to work with any model (model-agnostic) and to produce easily understandable results (explanations) \citep{ribeiro:2016}. Conceptually, LIME fits a simple (interpretable) model, the explainer model,  meant to capture the behavior of the (complex) black-box model in a local region around a prediction of interest. The  simple model then provides interpretable estimates for variables that most influenced the prediction made by the complex model.

\autoref{fig:concept-plot-good} provides a visualization of this conceptual understanding of LIME. The two plots show the predictions from a hypothetical black-box model plotted against the two predictor variables used to fit the hypothetical model. The diamond shaped points represent a prediction of interest. A Gower distance metric \citep{gower:1971} is used to define locality: the size of the points represent the inverse of the Gower distance as a reflection of the proximity to the prediction of interest. Here, an exponent of \Sexpr{good_gower_power} is used to emphasize interest in a very local region around the point of interest. A ridge regression model weighted by the proximity values is used as the explainer model with the black-box predictions as the response variable and standardized versions of Features 1 and 2 as predictor variables. Standardized features allow direct comparisons of the model coefficients. The explainer model is depicted by the black lines in the figure.

The plot on the left of \autoref{fig:concept-plot-good} shows the relationship between the black-box predictions and Feature 1. Here, the explainer model is plotted with Feature 2 fixed at the observed value of Feature 2 for the prediction of interest. The explainer model captures the relationship in the immediate neighborhood around the prediction of interest  with a slope of \Sexpr{coefs_data %>% filter(case == "good", feature == "Feature 1") %>% pull(slope) %>% round(3)}. The plot on the right shows that there is no global or local relationship between the black-box predictions and Feature 2. Here, the explainer model is plotted with Feature 1 set to be the observed value of Feature 1 for the prediction of interest, and it has an appropriately small slope of \Sexpr{coefs_data %>% filter(case == "good", feature == "Feature 2") %>% pull(slope) %>% round(3)}. The magnitude of the slope associated with Feature 1 is larger than the slope of Feature 2, which suggests that Feature 1 plays a more important role in the prediction made by the black-box model for the point of interest. This explanation agrees with a visual assessment of the relationships between the predictions and predictor variables.

@@

The concept of LIME is relatively simple: use an interpretable model to approximate a complex model in a local region. However, a practical implementation of LIME is not straightforward, and there is research is being done to improve the procedure \citep{laugel:2018}. The current implementations of LIME\footnote{https://github.com/thomasp85/lime} \footnote{https://github.com/goodekat/lime} offer various tuning parameters the user specifies when applying LIME (see Section \ref{background}) that affect the explainer model and ultimately, the explanation. Since the explainer model is an approximation of the complex model and not a direct interpretation, the explanations produced by an explainer model are subject to the quality of the approximation. Thus, in order to achieve accurate explanations, the tuning parameters selected need to be assessed.

We consider an example where the choice of exponent used to specify the weights is crucial to the quality of the explanation. The plots in Figure \ref{fig:concept-plot-bad} show the same data as \autoref{fig:concept-plot-good}, but the Gower distance metric exponent was decreased to \Sexpr{bad_gower_power} (the default exponent in the \emph{lime} R package). This causes the observations that are further away from the prediction of interest to be given larger weights than before. In \autoref{fig:concept-plot-good}, the explainer model captures the relationship between the black-box predictions in the immediate neighborhood of the prediction of interest. This cannot be said of the explainer model in \autoref{fig:concept-plot-bad}. In addition, the magnitude of the slope (\Sexpr{coefs_data %>% filter(case == "bad", feature == "Feature 2") %>% pull(slope) %>% round(3) %>% abs()}) associated with Feature 2 is larger than that of Feature 1 (\Sexpr{coefs_data %>% filter(case == "bad", feature == "Feature 1") %>% pull(slope) %>% round(3) %>% abs()}). Thus, this explainer model actually provides a misleading explanation that Feature 2 plays a more important role in predicting the observation of interest.

Several sources in the literature discuss the performance of LIME. One of the biggest difficulties with LIME is determining how to specify a local region \citep{molnar:2019} \citet{laugel:2018}. This is due to an unclear definition of what a local region means and how to apply LIME to achieve an appropriate local region as demonstrated by \autoref{fig:concept-plot-bad}. \citet{alvarezmelis:2018} raise a concern pertaining to the robustness of explanations from LIME and other model explainers: they find that even small changes in predictor variables can lead to very different LIME explanations. Additionally, \citet{ribeiro:2016} acknowledge that if a linear models is used as the explainer, LIME relies on a linear approximation of the explainer model to the complex model and state ``if the underlying model is highly non-linear even in the locality of the prediction, there may not be a faithful explanation".

As a result of the various ways LIME explanations can fail, it is important to assess LIME explanations. We suggest the use of visual diagnostics for assessment. In this paper, we lay out the set of claims about LIME made by \citet{ribeiro:2016} and propose four visualizations for the assessment of these claims: (1) \emph{explanation scatterplot}, (2) \emph{prediction comparison plot}, (3) \emph{feature heatmap}, (4) \emph{assessment metric plot}. While LIME has been implemented for image, tabular, and text data, we will only focus on tabular data. For additional simplicity, we will only discuss classification prediction models with a dichotomous response variable and continuous predictor variables. However, the proposed diagnostics may be extended to a wider range of situations.

\subsection{Structure of Paper}

The remainder of the paper is structured as follows. Section \ref{background} provides background and claims made by \citet{ribeiro:2016} about LIME. We introduce the suggested diagnostic plots in Section \ref{diagnostics}. Then in Section \ref{application}, we demonstrate the use of the diagnostics to assess LIME explanations for a random forest model fit to a forensics bullet matching dataset. Section \ref{discussion} concludes with a discussion on extension and limitations of the diagnostic plots and concerns about LIME in regards to the claims made by \citet{ribeiro:2016} brought about by the visualization examples in this paper that agree with \citet{alvarezmelis:2018}, \citet{laugel:2018}, and \citet{molnar:2019}.

@@

LIME was introduced in 2016 by \citet{ribeiro:2016}. The original authors provide an implementation of LIME in a Python package\footnote{https://github.com/marcotcr/lime}. An adaption of the Python package in R is by Thomas Lin-Pedersen\footnote{https://github.com/thomasp85/lime}. In this paper, any details provided about the implementation of LIME are based on the R package.

As described by \citet{laugel:2018}, the general form of the LIME algorithm can be divided into three steps:

@@

\item \emph{Explainer Model Fitting}: Apply the black-box model to the simulated data to obtain predictions. Compute the distance between each of the simulated observations and the prediction of interest. Perform feature selection. Fit an interpretable model with the black-box predictions from the simulated data as the response, the selected features from the transformed simulated data as the predictors, and the distances as weights. This model is the explainer model.

@@

During the application of LIME, the user must select various tuning parameter options: the number of features to return in the explanation, the simulation method, the feature selection method, and how the weights are computed. An overview of the options available for the tuning parameters is included in Appendix \ref{lime-details}.

In the original paper, \citet{ribeiro:2016} make the following set of claims regarding the performance of LIME:

@@

@@

In this section, we introduce four visual diagnostic plots for the assessment of LIME. The plots focus on different levels of application of LIME (e.g. on explanation versus a set of explanations) to assess the LIME claims from different perspectives:

\begin{enumerate}
\item \emph{Explanation Scatterplot} (Section \ref{plot1}): Comparison of the explainer and complex models for an individual prediction of interest.
\item \emph{Prediction Comparison Plot} (Section \ref{plot2}): Comparison of explainer and complex model predictions for a set of predictions.
\item \emph{Feature Heatmap} (Section \ref{plot3}): Comparison of features selected by LIME across applications of LIME with different tuning parameters.k
\item \emph{Assessment Metric Plot}: Comparison of performance metrics for LIME across applications of LIME with different tuning parameters. 
\end{enumerate}

@@

The dataset is divided into a training set of \Sexpr{dim(sine_data_train)[1]} observations and a testing set of \Sexpr{dim(sine_data_test)[1]} observations. A random forest model is fit using the R package \emph{randomForest} (version \Sexpr{packageVersion("randomForest")}) \citep{liaw:2002} with the default settings. The model is applied to the test set to obtain predictions. \autoref{fig:sine-plot-data} shows scatterplots of $x_2$ versus $x_1$ from the training data (left) and the testing data (right). Both plots include the true classification boundary of the rotated sine function plotted as the solid black line. The training data are colored by the observed response variable ($y$), and the testing data are colored by the prediction probabilities from the random forest model ($\hat{y}$). The random forest model misclassifies \Sexpr{dim(sine_data_test %>% filter(y != rfpred))[1]} points on the classification boundary. These are identified by circles in \autoref{fig:sine-plot-data}.

For the presentation of the explanation scatterplot, we focus on the misclassified point with $(x_1, x_2, x_3)$ coordinates of (\Sexpr{round(sine_poi$x1, 2)}, \Sexpr{round(sine_poi$x2, 2)}, \Sexpr{round(sine_poi$x3, 2)}) indicated by a diamond in \autoref{fig:sine-plot-data}. For the introduction of the other three plots, we use all observations in the \data test data as points of interest.

@@

\kgc{This is where I left off with the editing:} We apply LIME to the \data \ using \emph{lime} with the default settings. The user is asked to specify the number of features shown in the explanations. Here, we set this value to two, because we expect that $x_1$ and $x_2$ will be the main drivers in the prediction. Note that the complexity of an explanation goes hand in hand with the number of features selected to be included in the explanation. Setting the number of features to a small number will make the explanation simpler but not necessarily more correct.

To demonstrate the proposed diagnostic plots, we applied six different implementations of LIME to the sine data to select the top two features associated with each of the random forest predictions in the testing dataset. This was done using six different sets of tuning parameters via \sout{our} R package \emph{limeaid} \footnote{Available on GitHub at https://github.com/goodekat/limeaid}. \kge{(In this paper, we use a forked version\footnote{https://github.com/goodekat/lime} of the development version of the R package \emph{lime} (\Sexpr{packageVersion("lime")}) \citep{pedersen:2020}. The forked version of \emph{lime} allows us to export internal values of the procedure not already exported by \emph{lime}.)} Quantile bins were used for five of the implementations, and the number of bins for these implementations was varied from 2 to 6. The sixth implementation of LIME used the kernel density estimates. The feature selection method and the computation of the weights were set to the default options in \emph{lime} of forward selection and Gower distance raised to an exponent of 1.

@@

@@

\subsection{Feature Heatmap Diagnostic Plot} \label{plot3}

@@

\subsection{Assessment Metric Plot} \label{plot4}

@@

This paper highlights that while an explainer model is meant to provide clarity, it actually adds another layer of complexity to predictive models by requiring yet another model that needs to be assessed. Without an assessment of the explainer model, LIME is a black-box procedure of its own. We suggest the use of visual diagnostics to counteract the black-box nature of LIME and provide four diagnostic plots. Here we discuss the limitations and possible extensions of the diagnostic plots and concerns about LIME raise by applications of the diagnostics plots in this paper.

\subsection{Diagnostic Plot Limitations and Possible Extensions} \label{limit-exten}

% We acknowledge that not all prediction problems will produce trends as easy to interpret as the visualizations we have shown in the section, but we still believe that visualizations of the procedure can provide insight into the quality of the LIME explanation. We hope these figures will provide a starting point (if not an ending point) for the assessment of a single LIME explanation. Perhaps the inability to interpret a plot is itself telling, or the act of creating extensions of these visualizations to work with a different scenarios will lead to an understanding of the trustworthiness of the explanation.

%\hh{In terms of limitations of the visualizations: first two features span the x and y axis, additional features can be used for faceting. four features is completely feasible, beyond that the visualization might not work as well. The number of bins should be in the single digits. A large number of small multiples \citep{Becker:1996gy} needs a strong signal in the data to be effective - otherwise the visualization might be affected by change blindness \citep{simonsChangeBlindness1997, review19}. }

@@

@@

% Add a discussion on using our visualizations in situations without tabular data or a dichotomous response variable.

\subsection{Concerns about LIME} \label{concerns}

The visualizations are intended to provide insight on how LIME works, assess the ability of the explainer model to capture the complex predictive model, and compare LIME explanations produced by different tuning parameters. While the visualizations accomplish these tasks, they also expose examples of the failings of LIME. To address the discovered failings of LIME, we will reconsider each of the claims about the performance of LIME made by \citep{ribeiro:2016} (interpretability, faithfulness, linearity, and localness) in light of the insights gained from the diagnostic visualizations.

@@

Even though an explanation will be interpretable as long as the complexity of the explainer model is appropriately chosen, the presentation of a LIME explanation in the key visual used by \citep{ribeiro:2016} (shown in \autoref{fig:sine-plot-poi}) could lead to an over simplified interpretation of the explainer model. Without a solid understanding of the details behind the fitting of the explainer model, the deeper meaning of the bars in the figure is lost.

@@

@@

\section{LIME Tuning Parameter Options} \label{lime-details}
